Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States

### 2017-09-06 doc/RMD/MLG-distribution.Rmd
update figures

plot_mlg_graph <- function(g, glayout = NULL, label = TRUE, seed = 2017-06-28){
  set.seed(seed)
  if (is.null(glayout)){
    glayout <- layout_nicely(g)
    colnames(glayout) <- c("x", "y")
  } else {
    glayout <- glayout[V(g)$name, ]
  }
  shared_mlg  <- V(g)$weight * V(g)$size
  g           <- add_vertices(g, length(V(g)), size = V(g)$size - shared_mlg, color = "grey90")
  g2          <- g
  V(g2)$label <- ifelse(!is.na(V(g2)$name), sprintf("%s [(%d/%d)]", V(g2)$name, shared_mlg, V(g2)$size), NA)
  glay        <- create_layout(g2, "manual", node.positions = as.data.frame(rbind(glayout, glayout)))
  x_nudge     <- ifelse(abs(glay$x) == 1, -glay$x/10, glay$x/10)
  breaks      <- quantile(1 - E(g)$weight)
  breaks      <- setNames(breaks, format(breaks, digits = 2))
  outg <- ggraph(glay) +
    geom_edge_fan(aes(alpha = 1 - weight, width = width + 1)) +
    geom_node_circle(aes(r = drop(scale(size, center = FALSE))/10, fill = size))
  outg <- if (label) outg + geom_node_label(aes(label = label), repel = TRUE, parse = TRUE, label.size = 0.75, nudge_x = x_nudge) else outg
  outg + 
    viridis::scale_fill_viridis(option = "C") +
    scale_edge_alpha_continuous(range = c(1, 0.25), breaks = rev(breaks)) +
    scale_edge_width_continuous(range = c(0.25, 1.25), breaks = c(2:5)) +
    coord_fixed() +
    theme_void() +
    theme(text = element_text(size = 14)) +
    labs(list(
     # title = "Shared MLGs across regions",
      fill = "Number of\nMLHs",
      edge_alpha = "Probability of\nsecond encounter",
      edge_width = "Populations\nper MLH"#,
      #caption = "Outer circle: Number of MLGs in the region\nInner Circle: Number of private MLGs in the region"
    ))
}

>>>>

plot_mlg_graph <- function(g, glayout = NULL, label = TRUE, seed = 2017-06-28){
  set.seed(seed)
  if (is.null(glayout)){
    glayout <- layout_nicely(g)
    colnames(glayout) <- c("x", "y")
  } else {
    glayout <- glayout[V(g)$name, ]
  }
  shared_mlg  <- V(g)$weight * V(g)$size
  g           <- add_vertices(g, length(V(g)), size = V(g)$size - shared_mlg, color = "grey90")
  g2          <- g
  V(g2)$label <- ifelse(!is.na(V(g2)$name), sprintf("%s [(%d/%d)]", V(g2)$name, shared_mlg, V(g2)$size), NA)
  glay        <- create_layout(g2, "manual", node.positions = as.data.frame(rbind(glayout, glayout)))
  x_nudge     <- ifelse(abs(glay$x) == 1, -glay$x/10, glay$x/10)
  breaks      <- quantile(1 - E(g)$weight)
  breaks      <- setNames(breaks, format(breaks, digits = 2))
  outg <- ggraph(glay) +
    geom_edge_fan(aes(alpha = 1 - weight, width = width + 1)) +
    geom_node_circle(aes(r = drop(scale(size, center = FALSE))/10, fill = size))
  outg <- if (label) outg + geom_node_label(aes(label = label), repel = TRUE, parse = TRUE, label.size = 0.75, nudge_x = x_nudge, segment.size = 0, segment.alpha = 0) else outg
  outg + 
    viridis::scale_fill_viridis(option = "C") +
    scale_edge_alpha_continuous(range = c(1, 0.25), breaks = rev(breaks)) +
    scale_edge_width_continuous(range = c(0.25, 1.25), breaks = c(2:5)) +
    coord_fixed() +
    theme_void() +
    theme(text = element_text(size = 14)) +
    labs(list(
     # title = "Shared MLGs across regions",
      fill = "Number of\nMLHs",
      edge_alpha = "Probability of\nsecond encounter",
      edge_width = "Populations\nper MLH"#,
      #caption = "Outer circle: Number of MLGs in the region\nInner Circle: Number of private MLGs in the region"
    ))
}

### 2017-09-06 doc/RMD/by-year.Rmd
update figures

```{r packages}
library("poppr")
library("tidyverse")
library("cowplot")
make_amova_table <- function(am, amt, samples = "Region"){
  tot <- nrow(am$results)
  res <- data.frame(list(am$results[-tot, c("Df", "Sum Sq")], 
                         Percent = am$componentsofcovariance[-tot, 2],
                         Pval    = rev(amt$pvalue), 
                         Sigma   = am$componentsofcovariance[-tot, 1],
                         Phi     = rev(am$statphi$Phi[-tot])))
  res <- as.matrix(res)
  colnames(res) <- c("d.f.", "Sum of Squares", "Percent variation", "P", 
                     "Phi statistic")
  names(dimnames(res)) <- c("levels", "statistic")
  rownames(res) <- gsub("samples", samples, rownames(res))
  return(res)
}

@@

yearscale <- viridis::viridis(nlevels(strata(dat11cc)$Year), option = "C")
names(yearscale) <- sort(levels(strata(dat11cc)$Year))
gg_region_year <- ggscatter(regyear.dapc, strata(dat11cc)) +
  facet_wrap(~Region) +
  theme(legend.position = c(0.75, 0.1)) +
  guides(color = guide_legend(nrow = 4)) +
  scale_color_manual(values =  yearscale, breaks = names(yearscale)) +
  theme(legend.direction = "horizontal") +
  theme(strip.background = element_rect(color = NA, fill = "grey90")) +
  theme(strip.text = element_text(face = "bold", hjust = 0.05)) +
  theme(panel.border = element_blank())
gg_region_year
if (!interactive()) {
  ggsave(filename = file.path(PROJHOME, "results/figures/publication/dapc_region_year.pdf"),
         plot = gg_region_year,
         width = 7,
         height = 7)
}

>>>>

```{r packages}
library("poppr")
library("tidyverse")
library("cowplot")
make_amova_table <- function(am, amt, samples = "Region"){
  tot <- nrow(am$results)
  res <- data.frame(list(am$results[-tot, c("Df", "Sum Sq")], 
                         Percent = am$componentsofcovariance[-tot, 2],
                         Pval    = rev(amt$pvalue), 
                         Sigma   = am$componentsofcovariance[-tot, 1],
                         Phi     = rev(am$statphi$Phi[-tot])))
  res <- as.matrix(res)
  colnames(res) <- c("d.f.", "Sum of Squares", "Percent variation", "P", 
                     "Sigma", "Phi statistic")
  names(dimnames(res)) <- c("levels", "statistic")
  rownames(res) <- gsub("samples", samples, rownames(res))
  return(res)
}

@@

yearscale <- viridis::viridis(nlevels(strata(dat11cc)$Year), option = "C", end = 0.9)
names(yearscale) <- sort(levels(strata(dat11cc)$Year))
gg_region_year <- ggscatter(regyear.dapc, strata(dat11cc)) +
  facet_wrap(~Region) +
  theme(legend.position = c(0.75, 0.1)) +
  guides(color = guide_legend(nrow = 4)) +
  scale_color_manual(values =  yearscale, breaks = names(yearscale)) +
  theme(legend.direction = "horizontal") +
  theme(strip.background = element_rect(color = NA, fill = "grey90")) +
  theme(strip.text = element_text(face = "bold", hjust = 0.05)) +
  theme(panel.border = element_blank())
gg_region_year
if (!interactive()) {
  ggsave(filename = file.path(PROJHOME, "results/figures/publication/dapc_region_year.pdf"),
         plot = gg_region_year,
         width = 7,
         height = 7)
}

### 2017-09-06 doc/RMD/pop-diff.Rmd
update figures

```{r dapc-region-prob, fig.width = 7, fig.height = 6}
region_summary <- region.dapc$posterior %>% 
  as.data.frame() %>% 
  as_tibble() %>% 
  add_column(Individual = indNames(dat11cc)) %>% 
  add_column(`Original Population` = as.character(strata(dat11cc)$Region)) %>% 
  group_by(`Original Population`) %>%
  mutate(n = n()) %>%
  gather(Population, Posterior, -Individual, -`Original Population`, -n) %>% 
  group_by(Population, `Original Population`, n) %>% 
  summarize(`Mean Assignment` = mean(Posterior)) %>% 
  arrange(`Mean Assignment`) %>% 
  mutate(Source = case_when(
    Population == `Original Population` ~ "Original Population",
    TRUE ~ "Other"
    )) %>%
  ungroup() %>%
  arrange(desc(n)) %>%
  mutate(`Original Population` = forcats::fct_inorder(`Original Population`)) %>%
  mutate(Population = factor(Population, levels(`Original Population`))) %>%
  mutate(`Original Population` = paste0(`Original Population`, " (", n, ")")) %>%
  mutate(`Original Population` = forcats::fct_inorder(`Original Population`))
REGION_PLOT <- region_summary %>%
  filter(n >= 10) %>%
  ggplot(aes(x = Population, y = `Mean Assignment`)) +
  geom_segment(aes(xend = Population, yend = 0)) +
  geom_point(aes(fill = Source), pch = 21, size = 2) + 
  facet_wrap(~`Original Population`, ncol = 3, strip.position = "top") +
  scale_fill_grey() +
  # scale_x_discrete(position = "top") +
  scale_y_continuous(limits = c(0, 1), minor_breaks = c(seq(0, 1, by = 0.125)), breaks = c(0, 0.5, 1)) +
  theme_bw(base_size = 16, base_family = "Helvetica") +
  theme(legend.position = "top") +
  theme(aspect.ratio = 1/2) +
  theme(axis.text = element_text(color = "black")) +
  theme(axis.ticks.y = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  # theme(axis.text.x.top = element_text(angle = 90, vjust = 0.5, hjust = 0)) +
  theme(panel.grid.major.x = element_line(linetype = 0, color = "grey50")) +
  theme(panel.grid.major = element_line(colour = "grey20")) +
  theme(panel.grid.minor = element_line(linetype = 3, colour = "grey50")) +
  # theme(panel.spacing.y = unit(0, "line")) +
  theme(strip.background = element_rect(color = NA, fill = "grey90")) +
  theme(strip.text = element_text(face = "bold", hjust = 0.05)) +
  theme(panel.border = element_blank()) +
  xlab("Population Assignment") +
  ylab("Mean Probability")
REGION_PLOT

@@

```{r dapc-region-dualplot, fig.width = 13, fig.height = 6}
LEG <- cowplot::get_legend(LDS_PLOT + theme(legend.position = "right") + theme(legend.text = element_text(size = 12)))
cowplot::plot_grid(REGION_PLOT +
                     theme(plot.margin = unit(c(0, 0, 0, 0), "lines")), 
                   LDS_PLOT + 
                     theme(legend.position = "none") +
                     theme(plot.margin = unit(c(0, 0, 0, 0), "lines")) +
                     theme(aspect.ratio = 0.85), 
                   LEG,
                   nrow = 1, 
                   rel_widths = c(1, 0.85, 0.25),
                   label_size = 25,
                   label_fontfamily = "Helvetica",
                   labels = c("A", "B", NA),
                   align = "v",
                   axis = "t")
cowplot::ggsave(filename = file.path(PROJHOME, "results/figures/publication/DAPC.pdf"),
       width = 13,
       height = 6
      )
```

>>>>

```{r dapc-region-prob, fig.width = 7, fig.height = 6}
region_summary <- region.dapc$posterior %>% 
  as.data.frame() %>% 
  as_tibble() %>% 
  add_column(Individual = indNames(dat11cc)) %>% 
  add_column(`Original Population` = as.character(strata(dat11cc)$Region)) %>% 
  group_by(`Original Population`) %>%
  mutate(n = n()) %>%
  gather(Population, Posterior, -Individual, -`Original Population`, -n) %>% 
  group_by(Population, `Original Population`, n) %>% 
  summarize(`Mean Assignment` = mean(Posterior)) %>% 
  arrange(`Mean Assignment`) %>% 
  mutate(Source = case_when(
    Population == `Original Population` ~ "Original Population",
    TRUE ~ "Other"
    )) %>%
  ungroup() %>%
  arrange(desc(n)) %>%
  mutate(`Original Population` = forcats::fct_inorder(`Original Population`)) %>%
  mutate(Population = factor(Population, levels(`Original Population`))) %>%
  mutate(`Original Population` = paste0(`Original Population`, " (", n, ")")) %>%
  mutate(`Original Population` = forcats::fct_inorder(`Original Population`))
REGION_PLOT <- region_summary %>%
  filter(n >= 10) %>%
  ggplot(aes(x = Population, y = `Mean Assignment`)) +
  geom_segment(aes(xend = Population, yend = 0)) +
  geom_point(aes(fill = Source), pch = 21, size = 2) + 
  facet_wrap(~`Original Population`, ncol = 3, strip.position = "top") +
  scale_fill_grey() +
  # scale_x_discrete(position = "top") +
  scale_y_continuous(limits = c(0, 1), minor_breaks = c(seq(0, 1, by = 0.125)), breaks = c(0, 0.5, 1), position = "right") +
  theme_bw(base_size = 16, base_family = "Helvetica") +
  theme(legend.position = "top") +
  theme(aspect.ratio = 1/2) +
  theme(axis.text = element_text(color = "black")) +
  theme(axis.ticks.y = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  # theme(axis.text.x.top = element_text(angle = 90, vjust = 0.5, hjust = 0)) +
  theme(panel.grid.major.x = element_line(linetype = 0, color = "grey50")) +
  theme(panel.grid.major = element_line(colour = "grey20")) +
  theme(panel.grid.minor = element_line(linetype = 3, colour = "grey50")) +
  # theme(panel.spacing.y = unit(0, "line")) +
  theme(strip.background = element_rect(color = NA, fill = "grey90")) +
  theme(strip.text = element_text(face = "bold", hjust = 0.05)) +
  theme(panel.border = element_blank()) +
  xlab("Population Assignment") +
  ylab("Mean Probability")
REGION_PLOT

@@

```{r dapc-region-dualplot, fig.width = 13, fig.height = 6}
LEG <- cowplot::get_legend(LDS_PLOT + theme(legend.position = "right") + theme(legend.text = element_text(size = 12)))
cowplot::plot_grid(LEG,
                   LDS_PLOT + 
                     theme(legend.position = "none") +
                     theme(plot.margin = unit(c(0, 0, 0, 0), "lines")) +
                     theme(panel.border = element_rect(color = "grey50")) +
                     theme(aspect.ratio = 0.935), 
                   REGION_PLOT +
                     theme(plot.margin = unit(c(0, 1, 0, 0), "lines")), 
                   nrow = 1, 
                   rel_widths = c(0.25, 0.75, 1),
                   label_size = 25,
                   label_fontfamily = "Helvetica",
                   labels = c(NA, "A", "B"),
                   align = "v",
                   axis = "t")
cowplot::ggsave(filename = file.path(PROJHOME, "results/figures/publication/DAPC.pdf"),
       width = 13,
       height = 6
      )
```

### 2017-09-06 doc/manuscript/manuscript.Rmd
update figures

![
Discriminant Analysis of Principal Components on regions showing that Mexico is differentiated from other populations. 
**A)** Mean population assignment (horizontal axis) probability (vertical axis) for all populations with N > 10 (facets).
Numbers next to source populations indicate population size.
All values sum to one.
**B)** Scatter plot of first two components from DAPC.
Points represent observed individuals connected to the population centroids with ellipses representing a 66% confidence interval for a normal distribution. 
The center of each component is represented as black grid lines.
](../../results/figures/publication/DAPC.pdf){#DAPC width=100%}

>>>>

![
Discriminant Analysis of Principal Components on regions showing that Mexico is differentiated from other populations. 
**A)** Scatter plot of first two components from DAPC.
Points represent observed individuals connected to the population centroids with ellipses representing a 66% confidence interval for a normal distribution. 
The center of each component is represented as black grid lines.
**B)** Mean population assignment probability from the DAPC for all populations with N > 10 (facets).
Populations represented along the horizontal axis and probability of assignment on the vertical.
Numbers next to source populations indicate population size.
All values sum to one.
](../../results/figures/publication/DAPC.pdf){#DAPC width=100%}
